Databrary: A digital data library for sharing research video

### 2015-01-21 jcdl-15/paper.tex
some text and moving notes around

Video and audio recordings play a key role in many areas of research, especially in psychology, linguistics, education, ethnography, and ethology.
Video documents experimental procedures or behaviors, conveys ideas, provides examples for teaching and scientific communication, and serves as a rich primary data source for later analysis.
Unfortunately, in standard practice, the use of video data is restricted to the researchers who collect them, due largely to the perceived requirements around privacy restrictions.
The goal of the Databrary project is to create tools that allow video data to be documented, organized, and shared so that other researchers can discover and re-purpose them for their own research interests.
Accordingly, Databrary, with an initial focus on research in developmental psychology and related fields, has created a web-based library to accommodate the storage and organization of video data, capture contextual information necessary for re-use, and standardize across datasets without placing undue burden on data contributors.
In this paper, we discuss the design and technical challenges of the system and describe some findings and structural solutions to these challenges.
%rog The text you removed explained how and why video is actually "data." Maybe that's not important for this audience. What are the questions we are trying to answer with this paper? I'd argue for the following: 1) Why video? 2) What challenges does video pose? 3) How does Databrary solve these challenges? 4) What challenges remain unmet? 5) What lessons does Databrary offer other data repositories?
We have intentionally chosen to focus our efforts on hosting human-consumable, self-documenting, timeseries data: video and audio recordings.
Though while these types of data are an accessible and potent source for research analysis, they also present additional challenges \cite{Lanagan_Smeaton_2012}.
On one hand, recordings can be and largely are consumed and analyzed, at least initially, without sophisticated tools: once digitally decoded, trained human observers can glean directly much of the data's richness.
In particular, video data collected in one context can often be used by others for different purposes with minimal explanation, as the nature of the subjects and activity are readily apparent.
This naturally reduces the problems of documentation and re-use that plague most flat-file or dense datasets.
On the other hand, since automated analysis of videos is an active yet immature research area, the problems of search, discoverability, and meta-analysis over these media remain \cite{Albertson_2013}.
We attempt to address these issues by augmenting the primary recording data with additional structured and unstructured features.

>>>>

%rog The text you removed explained how and why video is actually "data." Maybe that's not important for this audience. What are the questions we are trying to answer with this paper? I'd argue for the following: 1) Why video? 2) What challenges does video pose? 3) How does Databrary solve these challenges? 4) What challenges remain unmet? 5) What lessons does Databrary offer other data repositories?

Video and audio recordings play a key role in many areas of research, especially in psychology, linguistics, education, ethnography, and ethology.
Video documents experimental procedures or behaviors, conveys ideas, provides examples for teaching and scientific communication, and serves as a rich primary data source for later analysis.
Unfortunately, in standard practice, the use of video data is restricted to the researchers who collect them, due largely to the perceived requirements around privacy restrictions.
The goal of the Databrary project is to create tools that allow video data to be documented, organized, and shared so that other researchers can discover and re-purpose them for their own research interests.
Accordingly, Databrary, with an initial focus on research in developmental psychology and related fields, has created a web-based library to accommodate the storage and organization of video data, capture contextual information necessary for re-use, and standardize across datasets without placing undue burden on data contributors.
In this paper, we discuss the design and technical challenges of the system and describe some findings and structural solutions to these challenges. The larger discussion involving research data repositories and the role of libraries in automating and facilitating access to research data well after its original use is very much active and ongoing \cite{Castelli_etal_2013, Nielson_Hjørland_2014, Macmillan_2014}. While Databrary is a model for tackling some of the challenges there, the scope of this paper focuses more on how we handle research data of a specific format and from a specific discipline, with implications for the future of data cataloging and reuse in other contexts.

We have intentionally chosen to focus our efforts on hosting human-consumable, self-documenting, timeseries data: video and audio recordings.
Though while these types of data are an accessible and potent source for research analysis, they also present additional challenges \cite{Lanagan_Smeaton_2012}.
On one hand, recordings can be and largely are consumed and analyzed, at least initially, without sophisticated tools: once digitally decoded, trained human observers can glean directly much of the data's richness.
In particular, video data collected in one context can often be used by others for different purposes with minimal explanation, as the nature of the subjects and activity are readily apparent.
This naturally reduces the problems of documentation and re-use that plague most flat-file or dense datasets.
On the other hand, since automated analysis of videos is an active yet immature research area, the problems of search, discoverability, and meta-analysis over these media remain \cite{Albertson_2013}.
We attempt to address these issues by augmenting the primary recording data with additional structured and unstructured features.
